---
output: github_document
---

<!-- README.md is generated from README.Rmd. Edit the .Rmd, then knit. -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width = 8,
  fig.height = 5.5,
  dpi = 110,
  message = FALSE,
  warning = FALSE
)
```

# lifr

<!-- badges: start -->
[![R-CMD-check](https://github.com/mjorden/lifr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/mjorden/lifr/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/mjorden/lifr/actions/workflows/pkgdown.yaml/badge.svg)](https://mjorden.github.io/lifr/)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![License: MIT](https://img.shields.io/badge/license-MIT-blue.svg)](LICENSE.md)
<!-- badges: end -->

**lifr** reads Laser-Induced Fluorescence (LIF) borehole logs from
high-resolution site characterization surveys (UVOST, TarGOST, and similar
direct-push fluorescence probes), lets you clean them with a recorded audit
trail, and turns them into plots and a site report.

| Stage | What lifr does |
|---|---|
| **Import** | Reads every `.lif.dat.txt` in a folder into one tidy data frame, keeping every channel the instrument recorded, and joins boring coordinates from a locations file. |
| **Edit** | Masks known-bad depth intervals. Every edit is logged in an audit trail that travels with the data and prints in the report. |
| **Summarise & plot** | QA statistics, depth profiles, multi-channel overviews, plan-view boring maps (optionally over USGS aerial imagery), depth-slice maps, and site summary charts. |
| **Report** | A self-contained HTML site report with Summary, Data & QA, Processing history, and Boring logs tabs, plus Markdown and PDF. |

Full documentation, with every function and worked articles, is at
<https://mjorden.github.io/lifr/>.

## Installation

```r
# install.packages("remotes")
remotes::install_github("mjorden/lifr")
```

lifr needs R 4.1 or newer. Optional extras: `ggrepel` for non-overlapping
map labels, `sf` and `png` for aerial-imagery basemaps, and `rmarkdown` plus
a LaTeX distribution (for example `tinytex::install_tinytex()`) for the PDF
report.

## Five-minute tour

The package ships with a synthetic 12-boring site so everything below runs
on a fresh install.

```{r import}
library(lifr)

demo <- system.file("extdata", "demo", package = "lifr")
lif <- lif_import(demo, verbose = FALSE)
lif
```

One row per depth reading. The known channels are canonicalised to `signal`
(%RE), `ec` (mS/m), `hp` (psi), and `color` (emission-wavelength hex); any
other column in the source files is carried through untouched.

### Look at the logs

```{r plot-all, fig.height = 6.5}
lif_plot_all(lif, ncol = 4)
```

```{r overview, fig.height = 4.5}
lif_overview(lif, "LIF-11")
```

### Clean up with an audit trail

Surface smear in the top foot and a rod-change artifact in one boring are
typical things to mask before summarising. Each call appends a row to the
edit history.

```{r edit}
lif <- lif_zero_shallow(lif, depth = 1)
lif <- lif_editor(lif, "LIF-03", top = 20, bottom = 22)
lif <- lif_keep(lif, "LIF-02", top = 6, bottom = 28)   # keep one clean window
tail(edit_history(lif), 3)
```

`qc_compare()` shows exactly what changed, with the edited depths in red:

```{r qc, fig.height = 4.5}
plots <- qc_compare(lif_import(demo, verbose = FALSE), lif)
plots[["LIF-02"]]
```

### Summarise

```{r summarize}
qa <- summarize_lif(lif)
qa$overview
head(qa$by_boring)
```

### Map the site

```{r map, fig.height = 7}
boring_map(lif, use_instrument_color = TRUE, site_name = "Demo site")
```

```{r chart, fig.height = 4}
chart_max_response(lif)
```

### Render the report

```r
site <- process_site(demo, output_dir = "out", site_name = "Demo site",
                     corrections = list(zero_shallow = 1))
site$reports$html
```

The HTML report is one self-contained file: stat tiles, the boring maps,
peak-response tables, QA statistics, the complete edit log, and a
one-boring-per-section log browser showing input beside edited data with
the zeroed windows shaded.

## Learn more

- `vignette("intro")`: getting started, end to end.
- `vignette("data-format")`: the raw file formats and what import does with them.
- `vignette("editing")`: masking artifacts, edit plans, and the audit trail.
- `vignette("plotting")`: a gallery of every plot with its options.
- `vignette("reporting")`: `process_site()`, the report tabs, and output files.

## About %RE

LIF response is relative fluorescence, expressed as a percentage of a
reference emitter (%RE). It indicates where fluorescent product is likely
present. It is not a concentration and is not comparable to regulatory
criteria; lifr never maps it onto such criteria.

## Synthetic data

The bundled demo site and every figure in the documentation come from
`lif_simulate()`. They have no relation to any real site. Use
`lif_simulate()` to try the package before you have field data.

## License

MIT. See `LICENSE.md`.
