---
title: "Raw files and what import does with them"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Raw files and what import does with them}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE)
```

```{r setup}
library(lifr)
```

## The survey folder

A LIF survey folder looks like this:

```
MySite/LIF/
├── LIF-01.lif.dat.txt
├── LIF-02.lif.dat.txt
├── ...
└── locations.csv
```

One log per boring, named for the boring, plus one locations table.

### Log files

A modern log is tab-delimited text with a header row. The first lines of a
demo file:

```{r head}
demo <- system.file("extdata", "demo", package = "lifr")
cat(readLines(file.path(demo, "LIF-05.lif.dat.txt"), n = 4), sep = "\n")
```

The exact set of columns depends on the probe and its firmware. `lifr`
recognises these headers (case and punctuation are ignored):

| Header seen in the field | Becomes | Notes |
|---|---|---|
| `Depth`, `Depth (ft)` | `depth` | Required |
| `Signal`, `LIF` | `signal` | Required |
| `EC.Value`, `EC Value`, `EC` | `ec` | Conductivity value (not `EC.Depth`, the channel's own offset depth) |
| `HP.PresDown`, `HP PresDown`, `HP` | `hp` | Downward hydraulic push pressure |
| `color`, `Color`, `colour` | `color` | Normalised to `#RRGGBB`; anything else becomes `NA` |

Every other header is kept under a cleaned snake_case name:

```{r read-one}
x <- lif_read(file.path(demo, "LIF-05.lif.dat.txt"))
names(x)
```

If a file carried, say, `EC.Depth` and `Detector 1 Max (uV)`, they would
come through as `ec_depth` and `detector_1_max_uv`. Nothing recorded by the
instrument is dropped at import.

Comma-separated `.lif.dat.csv` files are read the same way; the delimiter
is sniffed from the first line.

### Legacy files without a header

Older firmware wrote the same layout with no header row. Pass
`legacy = TRUE` and the columns are assigned by count: 11 columns is a LIF
log (`depth`, `signal`, `ch1`-`ch4`, `ec_depth`, `ec`, `hmr_depth`,
`hmr_val`, `color`) and 16 columns is a combined HP-LIF log (the same plus
`hp_depth`, `hp_time`, `hp_presup`, `hp`, `hp_flowup`).

```{r legacy, eval = FALSE}
lif <- lif_import("C:/Projects/OldSite/LIF", legacy = TRUE)
```

### The locations file

A CSV or tab-delimited table with one row per boring:

```{r locations}
read_locations(file.path(demo, "locations.csv"))
```

Column names are matched case-insensitively, and a few aliases are
accepted (`x`/`y` for easting/northing, `elevation` or `z` for `msl`,
`name` or `id` for `boring`). Coordinates should be in a projected system
in feet, typically the site's State Plane zone; `msl` is the ground-surface
elevation, used to convert depth to elevation.

A boring listed twice is an error. A duplicate row would double every
reading of that boring through the join, which is the kind of mistake that
is invisible in a plot and wrong in every statistic.

## What `lif_import()` does

1. Lists every file matching `.lif.dat.txt` or `.lif.dat.csv` in the
   folder and reads each with `lif_read()`.
2. Binds them. If the files do not agree on their column set (firmware
   drift, a hand-edited export), the union is used and the missing cells are
   `NA`, with a warning naming the files.
3. Finds the locations file (`locations.csv` is preferred over
   `locations.txt`, then any `locations*.csv`), or uses the path you give in
   `locations_file`.
4. Joins the coordinates on `boring`.
5. Orders the columns: `easting`, `northing`, `depth`, `signal`, `boring`,
   `msl`, `ec`, `hp`, `color`, then everything else.

```{r import}
lif <- lif_import(demo)
```

### When borings and locations disagree

Boring names come from the file names, so `LIF-01.lif.dat.txt` must match a
`LIF-01` row in the locations file exactly.

- A log with no locations row is **dropped with a warning that names it**.
  Fix the locations file and re-import.
- A locations row with no log is ignored, with a message.
- If **nothing** matches, import stops and prints both name lists side by
  side. That is almost always a format mismatch (`LIF-01` versus `LIF01`).

### Importing without coordinates

You can look at logs before the survey crew has sent coordinates:

```{r no-locs, eval = FALSE}
lif <- lif_import("C:/Projects/MySite/LIF", locations_file = FALSE)
```

Depth profiles and summaries work without coordinates. Maps and the report's
map section need them and will say so.

## The `lif_data` object

The import returns a plain data frame tagged with the `lif_data` class. The
class adds a compact `print()` and a `summary()` that calls
`summarize_lif()`; everything else you know about data frames, base R or
tidyverse, keeps working.

```{r class}
class(lif)
lif[lif$boring == "LIF-05" & lif$depth > 15 & lif$depth < 16, ]
```

To tag a frame you built some other way, use `new_lif_data()`; it checks for
`boring`, `depth`, and `signal`. `compat_rename_legacy()` maps older
TitleCase column names (`Depth`, `Signal`) onto the lowercase convention.

## Trying the package without field data

`lif_simulate()` writes a synthetic site in exactly this layout:

```{r simulate}
d <- tempfile("practice")
lif_simulate(n_borings = 6, dir = d, verbose = FALSE)
list.files(d)
sim <- lif_import(d, verbose = FALSE)
sim
```

The plume is a pair of Gaussian lobes, the pressure profile is hydrostatic
with thin low-pressure sand layers, and the emission colour is keyed to
signal strength. It is deliberately plausible and entirely made up.
