---
title: "Editing logs with an audit trail"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Editing logs with an audit trail}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7,
                      fig.height = 4.5, dpi = 72)
```

```{r setup, message = FALSE}
library(lifr)
demo <- system.file("extdata", "demo", package = "lifr")
lif <- lif_import(demo, verbose = FALSE)
```

## Why edit at all

LIF logs pick up things that are not product: smear in the top foot where
the probe was still settling, a spike at a rod change, a fluorescent mineral
seam, a stretch below total depth where the probe was already pulling back.
Left in, those readings inflate peak statistics and put false hot spots on
the map. Taken out silently, they leave a reviewer wondering what was
removed and why.

`lifr` editors do the second part properly. Every edit sets `signal` to a
replacement value (0 by default) over a depth window and appends a row to an
edit history that travels with the data frame and prints in the site report.

## The editor family

| Function | Zeroes | Scope |
|---|---|---|
| `lif_editor()` | inside a depth window | one boring |
| `lif_editor_bulk()` | inside a depth window | every boring |
| `lif_keep()` | **outside** a depth window | one boring |
| `lif_zero_shallow()` | at or above a depth | every boring or a subset |
| `lif_zero_below_threshold()` | wherever signal is below a value | every boring or a subset |
| `lif_apply_edits()` | whatever a plan table says | rows of the plan |

### Delete a window

```{r editor, message = TRUE}
lif <- lif_editor(lif, "LIF-03", top = 20, bottom = 22)
```

Several windows in one call:

```{r editor-multi, message = TRUE}
lif <- lif_editor(lif, "LIF-05", delete = list(c(0, 1.5), c(30, Inf)))
```

### Preview first

`preview = TRUE` reports how many readings would change and returns the
data unchanged:

```{r preview, message = TRUE}
lif_editor(lif, "LIF-07", top = 0, bottom = 3, preview = TRUE)
```

### Keep a window instead

When the honest description is "the clean interval is 6 to 28 ft", say that
directly rather than as two deletes:

```{r keep, message = TRUE}
lif <- lif_keep(lif, "LIF-02", top = 6, bottom = 28)
```

The audit row records the kept span and the zeroed rows are its complement.
The report's log plots read that note and shade the zeroed depths, not the
kept ones.

### Whole-site helpers

```{r helpers, message = TRUE}
lif <- lif_zero_shallow(lif, depth = 1)
lif <- lif_zero_below_threshold(lif, threshold = 0.5, borings = c("LIF-01", "LIF-04"))
```

## Edit plans in a CSV

For a real site, keep the edits in a file next to the data. One row per
edit:

```
boring,action,top,bottom
*,delete,0,1
LIF-03,delete,20,22
LIF-02,keep,6,28
LIF-05,delete,30,
```

- `boring`: a boring name, or `*` for every boring (delete rows only).
- `action`: `delete` (default), `clean` (an alias), or `keep`.
- `top`, `bottom`: blank means the function's own default, so `30,` reads
  "from 30 ft to the bottom of the boring".

Validate the plan before applying it. Every row is checked first, so a typo
in row 5 cannot half-apply rows 1 to 4:

```{r plan, message = TRUE}
plan <- data.frame(boring = c("*", "LIF-03", "LIF-02", "LIF-05"),
                   action = c("delete", "delete", "keep", "delete"),
                   top    = c(0, 20, 6, 30),
                   bottom = c(1, 22, 28, NA))
fresh <- lif_import(demo, verbose = FALSE)
lif_apply_edits(fresh, plan, validate = TRUE)
fresh <- lif_apply_edits(fresh, plan)
```

A `*` row that would span every reading at the site is refused; wiping the
whole dataset is never a one-liner.

## The audit trail

```{r history}
edit_history(fresh)
```

Each row carries a time-zone-stamped timestamp, the function, the boring
(`NA` means every boring), the window, the replacement value, how many rows
changed, and any notes. Save it beside the data for the project record:

```{r save, eval = FALSE}
edit_history_save(fresh, "C:/Projects/MySite/edit_log.csv")
```

Edited frames print a provenance banner so nobody mistakes them for raw
data:

```{r banner}
print(fresh, n = 2)
```

### Carrying the history through dplyr

dplyr verbs (`filter()`, `mutate()`, `select()`) rebuild the data frame and
drop non-standard attributes, so a pipe step silently discards the history.
Save it before and restore it after:

```{r rescue, eval = FALSE}
saved <- lif_get_edits(fresh)
sub   <- dplyr::filter(fresh, boring %in% c("LIF-02", "LIF-03"))
sub   <- lif_set_edits(sub, saved)
```

Base subsetting with `[` keeps the attribute, so `fresh[fresh$boring ==
"LIF-02", ]` needs no rescue.

## Corrections are edits too

`hp_correction()` removes the hydrostatic component of hydraulic push
pressure, `lif_downsample()` thins the depth spacing, and both record
themselves in the same history so the report can list them:

```{r corrections, message = TRUE}
fresh <- hp_correction(fresh, water_table = 6)
thin  <- lif_downsample(fresh, target_interval = 1, preserve_above = 5)
tail(edit_history(thin), 2)
```

## Seeing what changed

`qc_compare()` draws input beside edited for every boring that differs,
with the changed depths in red, and skips the rest:

```{r qc, message = TRUE}
plots <- qc_compare(lif_import(demo, verbose = FALSE), fresh)
names(plots)
plots[["LIF-05"]]
```

The site report does the same for every boring on its Boring logs tab,
with the zeroed windows shaded.
