---
title: "Introduction to lifr: from raw LIF logs to a site report"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to lifr: from raw LIF logs to a site report}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7,
                      fig.height = 5, dpi = 72)
```

Laser-Induced Fluorescence (LIF) probes such as UVOST and TarGOST are pushed
into the ground and record a fluorescence response every fraction of a foot.
Each boring produces one log file. This vignette walks through a complete
session with `lifr`: import the logs, look at the data, mask a couple of
artifacts with an audit trail, plot, and render the site report.

It uses the synthetic demo site that ships with the package, so every chunk
runs on a fresh install.

```{r setup}
library(lifr)
```

## 1. Import

A survey directory holds one `.lif.dat.txt` per boring plus a
`locations.csv` with each boring's easting, northing, and ground-surface
elevation. `lif_import()` reads everything, joins the coordinates, and keeps
every column the instrument recorded.

```{r import}
demo <- system.file("extdata", "demo", package = "lifr")
lif <- lif_import(demo)
lif
```

The known channels are canonicalised to `depth`, `signal`, `ec`, `hp`, and
`color`; any other column in the source file is carried through under a
cleaned name. `lif_read()` reads a single file the same way, without the
locations join.

## 2. Summarise

```{r summarize}
qa <- summarize_lif(lif)
qa$overview
head(qa$by_boring)
qa$by_threshold
```

`%RE` is relative fluorescence, a screening proxy for the presence of
fluorescent product. It is not a concentration, and `lifr` never compares it
to regulatory criteria.

## 3. Edit with an audit trail

Surface smear and known artifacts are masked with the editor family. Every
call appends a row to the frame's edit history.

```{r edit}
lif <- lif_zero_shallow(lif, depth = 1)
lif <- lif_editor(lif, "LIF-03", top = 20, bottom = 22)
lif <- lif_keep(lif, "LIF-02", top = 6, bottom = 28)
lif <- hp_correction(lif, water_table = 6)
tail(edit_history(lif), 4)
```

The same edits can live in a CSV and be applied with `lif_apply_edits()`;
a `boring` of `"*"` applies a row to every boring.

`qc_compare()` shows what changed:

```{r qc, fig.height = 4}
raw <- lif_import(demo, verbose = FALSE)
plots <- qc_compare(raw, lif)
plots[["LIF-02"]]
```

## 4. Plot

```{r plots}
lif_plot(lif, "LIF-11")
lif_plot_all(lif, ncol = 4)
lif_overview(lif, "LIF-11")
boring_map(lif, use_instrument_color = TRUE)
depth_slice_map(lif, depth_range = c(12, 20))
chart_max_response(lif)
```

With a projected CRS and network access, `fetch_basemap()` pulls a public
domain USGS imagery tile to draw under `boring_map()`.

## 5. Report

`process_site()` runs the whole pipeline in one call and writes a
self-contained HTML report (Summary, Data & QA, Processing history, and
Boring logs tabs), a Markdown report, and, when pandoc and LaTeX are
available, a PDF plus a one-page-per-boring logs PDF.

```{r report}
out <- tempfile("site")
site <- process_site(demo, output_dir = out, site_name = "Demo Site",
                     corrections = list(zero_shallow = 1, hp = list(water_table = 6)),
                     report_formats = c("html", "md"), quiet = TRUE)
site
basename(unlist(site$reports))
```

## 6. Try it on your own data

```{r own, eval = FALSE}
lif <- lif_import("C:/Projects/MySite/LIF")
lif <- lif_apply_edits(lif, "C:/Projects/MySite/edits.csv")
site <- process_site("C:/Projects/MySite/LIF",
                     edits = "C:/Projects/MySite/edits.csv",
                     output_dir = "C:/Projects/MySite/out",
                     crs = 3452)      # your State Plane EPSG code
```

No field data yet? `lif_simulate()` writes a plausible synthetic site.
