---
title: "Getting started with lifr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with lifr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7,
                      fig.height = 5, dpi = 72, message = FALSE)
```

## What LIF data looks like

A Laser-Induced Fluorescence (LIF) probe is pushed into the ground on a
direct-push rig. A laser excites the soil through a sapphire window and the
probe records the fluorescence response every fraction of a foot. Petroleum
products and other fluorescent non-aqueous phase liquids light up; clean soil
mostly does not. The result for each boring is a log: depth against
response, usually with a few companion channels such as electrical
conductivity, hydraulic push pressure, and an emission colour that hints at
the product type.

A survey delivers one log file per boring plus a table of boring
coordinates. `lifr` turns that folder into a data frame, helps you clean it,
and renders plots and a report.

```{r setup}
library(lifr)
```

## 1. Import

Point `lif_import()` at the survey folder. It reads every `.lif.dat.txt`,
finds the `locations.csv`, joins the coordinates, and returns one data frame
with a row per depth reading.

```{r import}
demo <- system.file("extdata", "demo", package = "lifr")
lif <- lif_import(demo)
lif
```

Every column the instrument recorded is kept. The ones `lifr` understands
are renamed to a fixed set:

| Column | Meaning | Units |
|---|---|---|
| `boring` | Boring name, from the file name | |
| `depth` | Depth below ground surface | ft |
| `signal` | LIF response | %RE |
| `easting`, `northing` | Coordinates from the locations file | ft (projected) |
| `msl` | Ground-surface elevation from the locations file | ft |
| `ec` | Electrical conductivity, when logged | mS/m |
| `hp` | Hydraulic push pressure, when logged | psi |
| `color` | Emission-wavelength hex colour, when logged | |

Anything else keeps a cleaned snake_case version of its header. See
`vignette("data-format")` for the file formats, legacy layouts, and what
happens when a boring is missing from the locations file.

## 2. Summarise

`summarize_lif()` returns the standard QA tables.

```{r summarize}
qa <- summarize_lif(lif)
qa$overview
head(qa$by_boring)
qa$by_threshold
```

A quick way to see which borings dominate the site:

```{r peaks, fig.height = 4}
chart_peak_by_boring(lif)
```

## 3. Clean up

Most sites need a little masking before the numbers mean anything: the top
foot or so is surface smear, and a rod change or a fluorescent mineral seam
can leave a spike that is not product. The editor family zeroes those depth
intervals and records every call.

```{r edit}
lif <- lif_zero_shallow(lif, depth = 1)                   # all borings
lif <- lif_editor(lif, "LIF-03", top = 20, bottom = 22)   # one interval
lif <- lif_keep(lif, "LIF-02", top = 6, bottom = 28)      # keep one clean window
tail(edit_history(lif), 3)
```

Compare before and after:

```{r qc, fig.height = 4.5}
raw <- lif_import(demo, verbose = FALSE)
plots <- qc_compare(raw, lif)
plots[["LIF-02"]]
```

The edit history travels with the data frame and is printed in the site
report. `vignette("editing")` covers edit plans in a CSV, the keep-versus-delete
choice, and how to carry the history through a dplyr pipeline.

## 4. Plot

```{r plot-one}
lif_plot(lif, "LIF-11")
```

```{r plot-all, fig.height = 6.5}
lif_plot_all(lif, ncol = 4)
```

```{r overview, fig.height = 4.5}
lif_overview(lif, "LIF-11")
```

```{r map, fig.height = 6.5}
boring_map(lif, use_instrument_color = TRUE, site_name = "Demo site")
```

`vignette("plotting")` is a gallery of every plot and its options,
including depth-slice maps, the site charts, and aerial-imagery basemaps.

## 5. Report

`process_site()` runs the whole pipeline in one call: import, an optional
edit plan, corrections, summary, charts, and the report.

```{r report}
out <- tempfile("site")
site <- process_site(demo, output_dir = out, site_name = "Demo site",
                     corrections = list(zero_shallow = 1),
                     report_formats = c("html", "md"), quiet = TRUE)
site
basename(unlist(site$reports))
```

The HTML report is one self-contained file with four tabs: Summary, Data &
QA, Processing history, and Boring logs. `vignette("reporting")` walks
through each tab and the other output files.

## 6. Your own data

```{r own, eval = FALSE}
lif  <- lif_import("C:/Projects/MySite/LIF")
lif  <- lif_apply_edits(lif, "C:/Projects/MySite/edits.csv")
site <- process_site("C:/Projects/MySite/LIF",
                     edits      = "C:/Projects/MySite/edits.csv",
                     output_dir = "C:/Projects/MySite/out",
                     site_name  = "My Site",
                     crs        = 3452)   # State Plane EPSG code, for imagery
```

No field data yet? `lif_simulate()` writes a plausible synthetic site:

```{r simulate, eval = FALSE}
lif_simulate(n_borings = 8, dir = "C:/Projects/Practice/LIF")
```

## A note on %RE

LIF response is relative fluorescence, expressed as a percentage of a
reference emitter. It tells you where fluorescent product is likely present
and how strongly it fluoresces there. It is not a concentration and is not
comparable to regulatory criteria; `lifr` never maps it onto such criteria,
and every chart and report carries that disclosure.
