---
title: "Plot gallery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plot gallery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7,
                      fig.height = 5, dpi = 72, message = FALSE, warning = FALSE)
```

```{r setup}
library(lifr)
library(ggplot2)
demo <- system.file("extdata", "demo", package = "lifr")
lif <- lif_import(demo, verbose = FALSE)
lif <- lif_zero_shallow(lif, depth = 1)
```

Every plotting function returns a ggplot (or a patchwork of them), so you
can add layers, change labels, or save with `ggsave()` as usual. All of them
share `theme_lifr()`.

## Depth profiles

### One boring: `lif_plot()`

Depth runs down the page. When the log carries an emission colour, the
trace is filled with it and a colour strip sits left of zero; otherwise a
single fill is used.

```{r lif-plot}
lif_plot(lif, "LIF-11")
```

Fix the signal axis to compare borings on the same scale, or cap the depth:

```{r lif-plot-opts, fig.height = 4}
lif_plot(lif, "LIF-05", ymax = 100, xmax = 25)
```

### Every boring: `lif_plot_all()`

```{r plot-all, fig.height = 6.5}
lif_plot_all(lif, ncol = 4)
```

`layout = "list"` returns one full `lif_plot()` per boring instead, and
`output_file` writes the facet figure straight to a PNG.

### Multi-channel: `lif_overview()`

LIF beside conductivity and pressure on a shared depth axis. Channels the
boring did not record are skipped.

```{r overview, fig.height = 4.5}
lif_overview(lif, "LIF-06")
```

### Input versus edited: `qc_compare()`

```{r qc, fig.height = 4.5}
edited <- lif_keep(lif, "LIF-10", top = 8, bottom = 24)
qc_compare(lif, edited)[["LIF-10"]]
```

## Plan-view maps

### `boring_map()`

Each boring is a marker sized by the square root of its peak signal, so the
site's pattern reads even in greyscale, and labelled with its name and
peak. Non-detects are hollow circles. A north arrow and scale bar sit in
whichever corners overlap the fewest borings.

```{r map, fig.height = 6.5}
boring_map(lif, site_name = "Demo site", figure_date = "2026-09-04",
           preparer = "Field team")
```

`use_instrument_color = TRUE` fills each marker with the emission colour at
the depth of peak signal, the product signature analysts read directly:

```{r map-color, fig.height = 6.5}
boring_map(lif, use_instrument_color = TRUE)
```

Restrict the peak to a depth window:

```{r map-window, fig.height = 6.5}
boring_map(lif, depth_min = 12, depth_max = 20, use_instrument_color = TRUE)
```

Other useful arguments: `nd_threshold` (peaks at or below it are
non-detects), `size_by_signal = FALSE` for uniform markers, `coord_units`
for metric sites, `crs_label` for the caption, `repel = FALSE` to skip
ggrepel, and `output_file` to save.

### Aerial imagery under the map

With a projected CRS and network access, `fetch_basemap()` pulls a
public-domain USGS National Map tile for the site (US coverage only) and
`boring_map()` draws it under the markers; labels switch to white with a
dark halo.

```{r basemap, eval = FALSE}
bm <- fetch_basemap(lif, crs = 3452)   # NAD83 / Louisiana South (ftUS)
boring_map(lif, basemap = bm)
```

`fetch_basemap()` needs the `sf` and `png` packages and returns `NULL` with
a warning when it cannot fetch, so a map call with `basemap = NULL` simply
draws without imagery.

### `depth_slice_map()`

Colour each boring by the aggregate (default maximum) of a channel inside a
depth band. Borings with no readings in the band are grey crosses, so "not
sampled here" is never mistaken for "no boring here".

```{r slice}
depth_slice_map(lif, depth_range = c(12, 20))
```

```{r slice-ec}
depth_slice_map(lif, depth_range = c(5, 10), channel = "ec", agg_fn = median)
```

## Site charts

These are the figures the site report embeds; each is also available on
its own.

```{r chart-max, fig.height = 4}
chart_max_response(lif, thresholds = c(1, 10, 50))
```

```{r chart-peak, fig.height = 4}
chart_peak_by_boring(lif)
```

```{r chart-hp, fig.height = 3.5}
chart_hp_histogram(lif)
```

```{r chart-ec, fig.height = 3.5}
chart_ec_histogram(lif)
```

`export_site_charts()` writes whichever of them the data supports as PNGs
in one call.

## Customising

Everything is a ggplot. Add layers or override labels:

```{r custom, fig.height = 4.5}
lif_plot(lif, "LIF-11") +
  geom_hline(yintercept = 10, linetype = "dashed", colour = "grey40") +
  labs(title = "LIF-11 with a 10 %RE screening line",
       subtitle = "Demo site, September 2026")
```

Use the package theme on your own figures:

```{r theme, fig.height = 4}
re <- re_feet(lif)
ggplot(re, aes(x = reorder(boring, re_feet), y = re_feet)) +
  geom_col(fill = "#6E7F4E") +
  coord_flip() +
  labs(x = NULL, y = "RE-feet (integrated signal)",
       title = "Integrated LIF response by boring") +
  theme_lifr()
```

Save any plot with `ggplot2::ggsave()`:

```{r save, eval = FALSE}
ggsave("boring_map.png", boring_map(lif), width = 8, height = 7, dpi = 200)
```
