---
title: "The site pipeline and report"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The site pipeline and report}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", message = FALSE)
```

```{r setup}
library(lifr)
demo <- system.file("extdata", "demo", package = "lifr")
```

## One call

`process_site()` chains the pipeline: import, an optional edit plan,
corrections, QA summary, site charts, and the report.

```{r run}
out  <- tempfile("site")
plan <- data.frame(boring = c("*", "LIF-02"), action = c("delete", "keep"),
                   top = c(0, 6), bottom = c(1, 28))

site <- process_site(
  demo,
  edits          = plan,                       # or a path to edits.csv
  corrections    = list(hp = list(water_table = 6)),
  output_dir     = out,
  site_name      = "Demo site",
  report_formats = c("html", "md"),
  quiet          = TRUE)
site
```

The returned `lif_site` object holds everything the report was built from:

```{r object}
names(site)
site$log
head(site$edits)
basename(unlist(site$reports))
```

`site$data` is the processed frame (classed, with its edit history) and
`site$raw_data` is the frame exactly as imported, so you can keep working
with either.

### Arguments worth knowing

| Argument | What it does |
|---|---|
| `edits` | A CSV path or data frame for `lif_apply_edits()`. See `vignette("editing")`. |
| `corrections` | `list(zero_shallow = 1)` zeroes the top foot; `zero_below = 0.5` zeroes readings under 0.5 %RE; `hp = TRUE` or `hp = list(water_table = 6, gradient = 0.433)` runs `hp_correction()`. Applied in that order, after the edits. |
| `crs` | The projected EPSG code of the coordinates. When given, the report adds a boring map over USGS aerial imagery (needs network access; degrades to a note offline). |
| `response_thresholds` | The Summary tab lists borings whose peak exceeds each value (default 1, 10, 50 %RE). |
| `report_formats` | Any of `"html"`, `"md"`, `"pdf"`. |
| `date_suffix` | Appends `_YYMMDD` to every output name so successive runs sit side by side. |
| `prefix` | Filename prefix; default is a slug of `site_name`. |

## What gets written

```{r files}
list.files(out)
```

- `<prefix>_report.html`: the self-contained HTML report.
- `<prefix>_report.md` plus `<prefix>_map_*.png` and `<prefix>_chart_*.png`
  sidecars: the Markdown report. Keep them together when moving it.
- `<prefix>_chart_*.png`: the site charts from `export_site_charts()`.
- With `"pdf"`: `<prefix>_report.pdf` (rendered from the Markdown through
  pandoc and LaTeX) and `<prefix>_logs.pdf`, one page per boring, drawn by
  R's own PDF device so it needs no LaTeX.

## The HTML report

Open the HTML file in any browser. It needs no server and no internet
connection; the figures are embedded and the tables sort when you click a
heading. Printing lays every tab out as its own section.

**Summary.** Stat tiles (borings, samples, depth range, channels, edits),
the boring maps, the %RE disclosure, a table of every boring's peak
response and the depth it occurred at, and one table per screening
threshold listing the borings above it.

**Data & QA.** A per-boring inventory (samples, top and bottom depth,
native sampling interval, channels present), the summary statistics from
`summarize_lif()` with units, the site charts, and a definitions footnote.

**Processing history.** The chronological record: what was imported, the
complete edit log, the corrections, and the pipeline steps with row and
boring counts. This is the audit trail a reviewer reads to answer "what did
you do to the data?"

**Boring logs.** One section per boring showing the input log beside the
edited log with the zeroed depth windows shaded red; whole-log operations
without a window are named beneath the panel. A "changed borings only"
filter and a navigation strip make a 40-boring site quick to scan, and each
changed boring carries a badge naming its edited windows.

```{r html}
html <- readLines(site$reports$html, warn = FALSE)
grep("data-tab=", html, value = TRUE)[1:4]
```

## Re-rendering without re-importing

`site_report()` renders any subset of formats from the object:

```{r rerender}
site_report(site, formats = "md", output_dir = out, file_prefix = "demo_again",
            quiet = TRUE)
```

That is the way to add a PDF after the fact, or to render into a different
folder.

## The PDF toolchain

The main PDF is produced by `rmarkdown` from the Markdown report, which
needs pandoc (bundled with RStudio) and a LaTeX distribution with
`xelatex`. The easiest way to get one:

```{r tinytex, eval = FALSE}
install.packages("tinytex")
tinytex::install_tinytex()
```

When either piece is missing, `process_site()` still writes the other
formats and the logs PDF, and says so in a warning.

## Reproducibility

Keep the survey folder, the edits CSV, and the `process_site()` call in a
script. Re-running it reproduces every figure and table, and the edit log in
the report records the exact windows applied. With `date_suffix = TRUE`,
successive runs are kept side by side instead of overwriting each other.
